---
title: "Assignment_5"
author: "Keith Rafferty"
date: "2024-09-29"
output: html_document
---
## Data Import, Tidying, and Transformation

The data is imported as a CVS file and converted to a data frame. The third row is dropped from the data frame because it is an empty row that served as a spacer between the data for the individual airlines in the CSV file. Next, the airline names are duplicated to fill in the missing values in rows 2 and 4 of column "X", which is then renamed "Airline". The data is then pivoted vertically so that the individual airport results are captured in a column containing the airport name and another column capturing the delay or on-time flight count. Lastly, two new columns are added to the data frame via the mutate() function: 1) the total flights at each airport for each air line and 2) the event rate (delay or on-time) for each airline at each airport. The completed tidy data set is displayed via kable() for clarity.

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)

#Load libraries needed for assignment.

library(tidyverse)
library(ggthemes)
library(dplyr)
library('RCurl')
library(tidyverse)
library(ggplot2)
library(knitr)
library(ggpubr
        )
#Import data set from GitHub

getdata <- getURL('https://raw.githubusercontent.com/kr0710/Data607/refs/heads/main/Assignment_5_data.csv')

#Convert data to a data frame, drop empty rows, fill in missing row and column names

df <- read.csv(text = getdata)[-c(3),]
df <- read.csv('c:/data/Assignment_5_data.csv')[-c(3),]
df$X <- c(df$X[1], df$X[1], df$X[3], df$X[3])
colnames(df)[1] <- "Airline"
colnames(df)[2] <- "flight_fate"

#Pivot dat vertically 

df_pivot <-
  pivot_longer(df,
    cols = colnames(df)[3:7],
    names_to = 'Airport',
    values_to = 'fate_count',
  )

#Add columns to calculate total flights at each airport for each airline
#Calculate delayed or on-time flight rates
df_final <-
  df_pivot |>
  group_by(Airline, Airport) |>
  mutate(total_flights = sum(fate_count)) |>
  mutate(fate_rate = round((fate_count/total_flights)*100, digits = 2))

#display tidy and transformed data
kable(df_final)

```
# Data Visualization and Analysis

The data frame is then bifurcated into two data frames depending on the status of the flight (i.e. "flight_fate"). The  "fate_rate" (i.e. rate of delay or on-time flights) of the airports and airlines in each respective data frame are both visualized via ggplot2.

At all five airports, it appears that Alaskan Airlines outperforms AM West in respect to lower rates of delayed flights. Naturally, Alaskan airlines also has higher on-time rates across the board at each airport surveyed. Phoenix airport is the best performing airport since it has the lowest delay rates and the highest on-time rates for each airline. Conversely, for both airlines, San Francisco airport is the worst performing airport in terms of delay and on-time rates.

The relationship between total flights and the delayed flight rate is examined in graphical form. However, there is not a strongly correlated relationship between these two variables, but obviously the data is quite limited (n=10). 

```{r echo = FALSE, results= 'asis'}
#Filter df_final to isolate delayed flight counts

delayed_df <-
  df_final |>
  filter (flight_fate == 'delayed') |>
  select(Airline, Airport, flight_fate, total_flights, fate_rate)

#Filter df_final to isolate on-time counts
on_time_df <-
  df_final |>
  filter (flight_fate == 'on time') |>
  select(Airline, Airport, flight_fate, total_flights, fate_rate)

# Display delayed flight data and graph

kable(delayed_df)

ggplot(delayed_df, aes(x = Airport, y = fate_rate, fill = Airline)) +
  geom_bar(position="dodge", stat="identity") +
  labs(y = "Delayed Flights (%)", title = "Delayed Flight Rates")

# Display on-time flight data and graph
kable(on_time_df)

ggplot(on_time_df, aes(x = Airport, y = fate_rate, fill = Airline)) +
  geom_bar(position="dodge", stat="identity") +
  labs(y = "On Time Flights (%)", title = "On Time Flight Rates")

#Examine the relationship b/t total flights at an airport and delay rates

ggplot(delayed_df, aes(x = total_flights, y = fate_rate)) +
  geom_point(aes(color = Airport)) +
  geom_smooth(method = "lm", se=FALSE, formula = y ~ x) +
  labs(y = "Delayed Flights (%)", x = "Total Flights",
  title = "Delayed Flight Rates vs. Total Flights") +
  stat_regline_equation(label.x=4100, label.y=30) +
  stat_cor(aes(label=after_stat(rr.label)), label.x=4100, label.y=28)

```

